HCM SCD risk algorithm cost-effectiveness analysis

### 2021-07-21 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
added all the components for the paper but the results are not final; they need to be run properly

---
title: "HCM SCD risk algorithm cost-effectiveness analysis"
author: "Nathan Green"
date: "09/11/2020"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
bibliography: bibliography.bib
---

@@

@@

There have been previous cost-effectiveness analyses [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]

@@

The main data set contains individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.

@@

+----------------------------------+-----------------+--------+------------------+
| Description                      | Value$^*$       | Range  | Source           |
+==================================+=================+========+==================+
| *Health*                         |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Manage with ICD                  | 0.637 QALY/year |        | @Noyes2007       |
+----------------------------------+-----------------+--------+------------------+
| Implantation procedure utility   | -0.016          |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Shock utility                    | -0.5            |        |                  |
+----------------------------------+-----------------+--------+------------------+
| HCM without ICD                  | 1 QALY/year     |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Death                            | 0 QALY/year     |        |                  |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| *Cost*                           |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| ICD appointment                  | £10             |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Perform risk score               | £20?            |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant ICD                      | £4,666          |        | EY02B Tariffs    |
+----------------------------------+-----------------+--------+------------------+
| Implant replacement              | £45,000         |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant complication             | £28,839         |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Non-fatal shock                  | £22,880         |        | UK Stroke Assoc. |
+----------------------------------+-----------------+--------+------------------+
| HCM without ICD                  | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
| SCD                              | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
| All-cause death                  | 0               |        |                  |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| *Probabilities*                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Initial implant complication     | 0.047           |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
| Replacement implant complication | 0.032           |        | @Smith2013       |
+----------------------------------+-----------------+--------+------------------+
|                                  |                 |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Time horizon                     | 12 years        |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Implant replacement              | 10 years        |        |                  |
+----------------------------------+-----------------+--------+------------------+
| Annual number of appointments    | 2               |        |                  |
+----------------------------------+-----------------+--------+------------------+

@@

@@

-   Partition observed in data set
-   ICD given if number of risk factors \> 0, 1 or 2
-   ICD given if risk score \>6%
-   ICD given if risk score \>4%

@@

A diagram of the current cohort model is given below.

```{r fig.align="center", echo=FALSE}
knitr::include_graphics("model_diagram.png")
```

Assuming that shocked patients return to the HCM ICD state then the transition matrix may look like the following.

@@

We assume that an ICD patient has 2 annual appointments.
All shocks are treated the same in terms of costs and health impact.
Implantation can have complications.
The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from [@Smith2013].

-   Annual health: $$
    e^0_{s=1}(t) = e^1_{s=1}(t) = u\_hcm + u\_icd
    $$

$$
e^0_{s=2}(t) = e^1_{s=2}(t) = u\_shock
$$

$$
e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 3,4,5,6
$$

-   Initial costs: $$
    c^0_{s=1}(t = 0) = c\_icd + p\_compl \times c\_compl
    $$

$$
c^1_{s=1}(t = 0) = c\_icd + c\_rscore + p\_compl \times c\_compl
$$

-   Annual cost: $$
    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 c\_icd\_appt, \;\; t \neq t\_repl
    $$

$$
c^0_{s=2}(t) = c^1_{s=2}(t) = c\_nfatal
$$

-   Implant replacement cost $$
    c^0_{s=1}(t = t\_repl) = c^1_{s=1}(t = t\_repl) = 2 c\_icd\_appt + c\_icd\_repl + p\_compl \times c\_compl
    $$

@@

$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$ $$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$

$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$ $$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$

@@

## Workflow

The main files to perform the analysis are:

-   `prep_study_data.R` munges the raw data and save the input data in `data/`
-   `BUGS/script.R` runs the BUGS code in `BUGS/model.txt`
-   `main-ce-analysis.R` performs the cost-effectiveness analysis
-   `pop_counts_plot.R` creates output plots

@@

We give some example output to demonstrate what will be produced with the final model.

Histogram of posterior distributions for state transition probabilities.

![](posterior_histograms.png)

Below is an example of state occupancy over time plot.
This shows that for the new algorithm there are fewer scd and more shocks.

![](../../images/state_pop_over_time.png)

>>>>

---
title: "HCM SCD risk algorithm cost-effectiveness analysis"
author: "Nathan Green, Yang Chen"
date: "09/11/2020"
output: 
  bookdown::pdf_document2:
    toc: false
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: sentence
params:
  data: "ipd_risk"
bibliography: bibliography.bib
---

@@

```{r, echo=FALSE}
data(list = params$data, package = "HCM.SCD.CEanalysis")
```

```{r echo=FALSE}
centre_tab <- dplyr::count(ipd_risk, centre)
centre <- centre_tab$centre
n <- centre_tab$n

cens_tab <- dplyr::count(ipd_risk, d)
cens <- cens_tab$d
nd <- cens_tab$n
```

@@

There have been previous cost-effectiveness analyses of ICD implants [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]

Cost-effectiveness of a cardiovascular risk prediction algorithm.
[@Zomer2017] for statins interventions and people with severe mental illness.

[@Yao2007] use a Markov model and consider multiple implantations attempts if unsuccessful.
[@Colquitt2014] is a Health Technology Assessment which compares optimal pharmacological therapy (OPT) with or without ICD.
[@Tomini2016] is a review of economic evaluation models for cardiac resynchronization therapy with implantable cardioverter defibrillators in patients with heart failure.
[@Ommen2020] 2020 AHA/ACC Guideline for the Diagnosis and Treatment of Patients With Hypertrophic Cardiomyopathy.

@@

The main data set contains individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.
Key cohort characteristics include the following.
Patients were enrolled from the 6 health centres: `r paste0(centre[1], " (", n[1], ")")`, `r paste0(centre[2], " (", n[2], ")")`, `r paste0(centre[3], " (", n[3], ")")`, `r paste0(centre[4], " (", n[4], ")")`, `r paste0(centre[5], " (", n[5], ")")`, `r paste0(centre[6], " (", n[6], ")")`; The amount of censoring was `r paste0(cens[1], " (", nd[1], ")")`, `r paste0(cens[2], " (", nd[2], ")")`; The mean age (sd) was `r round(mean(ipd_risk$age),0)` (`r round(sd(ipd_risk$age),0)`); The start of study data collection was in `r min(ipd_risk$year)` to `r max(ipd_risk$year)`. Further plots are given in the Appendix. See [@OMahony2014] for further details.

@@

+----------------------------------+-------------+-----------------+-------+------------------+
| Description                      | Parameter   | Value$^*$       | Range | Source           |
+==================================+=============+=================+=======+==================+
| *Health*                         |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Manage with ICD                  | `u_icd`     | 0.637 QALY/year |       | @Noyes2007       |
+----------------------------------+-------------+-----------------+-------+------------------+
| Implantation procedure utility   | `u_implant` | -0.016          |       | @Smith2013       |
+----------------------------------+-------------+-----------------+-------+------------------+
| Shock utility                    | `u_shock`   | -0.5            |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| HCM without ICD                  | `u_hcm`     | 1 QALY/year     |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Death                            | `u_death`   | 0 QALY/year     |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
|                                  |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| *Cost*                           |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| ICD appointment                  | `c_appt`    | £10             |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Perform risk score               | `c_rs`      | £20?            |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Implant ICD                      | `c_implant` | £4,666          |       | EY02B Tariffs    |
+----------------------------------+-------------+-----------------+-------+------------------+
| Implant replacement              | `c_repl`    | £45,000         |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Implant complication             | `c_compl`   | £28,839         |       | @Smith2013       |
+----------------------------------+-------------+-----------------+-------+------------------+
| Non-fatal shock                  | `c_shock`   | £22,880         |       | UK Stroke Assoc. |
+----------------------------------+-------------+-----------------+-------+------------------+
| HCM without ICD                  | `c_hcm`     | 0               |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| SCD                              | `c_scd`     | 0               |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| All-cause death                  | `c_death`   | 0               |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
|                                  |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| *Probabilities*                  |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Initial implant complication     | `p_compl`   | 0.047           |       | @Smith2013       |
+----------------------------------+-------------+-----------------+-------+------------------+
| Replacement implant complication | `p_repl`    | 0.032           |       | @Smith2013       |
+----------------------------------+-------------+-----------------+-------+------------------+
|                                  |             |                 |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Time horizon                     | `T`         | 12 years        |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Implant replacement              | `t_repl`    | 10 years        |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+
| Annual number of appointments    | n_appt      | 2               |       |                  |
+----------------------------------+-------------+-----------------+-------+------------------+

@@

: Starting state populations by decision rule.

@@

-   Partition observed in data set
-   ICD given if number of risk factors \> 0, 1 or 2
-   ICD given if risk score \> 6%
-   ICD given if risk score \> 4%

@@

A diagram of the current cohort model is given in Figure \@ref(fig:model).

```{r model, fig.align="center", echo=FALSE, fig.cap="Markov model diagram. Bold circles represent starting states and dashed circles represent sink states."}
knitr::include_graphics("../../images/model_diagram.png")
```

Assuming that shocked patients return to the HCM ICD state then the transition matrix looks like the following.

@@

We assume that an ICD patient has 2 annual appointments.
All shocks are treated the same in terms of costs and health impact.
Implantation can have complications.
The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from [@Smith2013].
Subscript $s$ denotes the state number and superscript denotes the intervention either ICD or not.

-   Annual health: $$
    e^0_{s=1}(t) = e^1_{s=1}(t) = \mbox{u\_hcm} + \mbox{u\_icd}
    $$ $$
    e^0_{s=3}(t) = e^1_{s=3}(t) = \mbox{u\_hcm}
    $$ $$
    e^0_{s=2}(t) = e^1_{s=2}(t) = \mbox{u\_shock}
    $$

$$
e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 4,5,6
$$

-   Initial costs: $$
    c^0_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{p\_compl} \times \mbox{c\_compl}
    $$

$$
c^1_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{c\_rscore} + \mbox{p\_compl} \times \mbox{c\_compl}
$$

-   Annual cost: $$
    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt}, \;\; t \neq \mbox{t\_repl}
    $$

$$
c^0_{s=2}(t) = c^1_{s=2}(t) = \mbox{c\_shock}
$$

-   Implant replacement cost $$
    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt} + \mbox{c\_repl} + \mbox{p\_compl} \times \mbox{c\_compl}, \;\;  t = \mbox{t\_repl}
    $$

@@

$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$

$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$

$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$

$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$

@@

@@

We give results of the model fitting and cost-effectiveness analysis.

### Model fitting

Figure \@ref(fig:hist) gives a histogram of posterior distributions for state transition probabilities.

```{r hist, fig.align="center", echo=FALSE, out.width = "80%", fig.cap="Histogram of posterior distributions for state transition probabilities."}
knitr::include_graphics("posterior_histograms.png")
```

Figure \@ref(fig:statepop) gives an example of state occupancy over time plot.
This shows that for the new algorithm there are fewer SCD and more shocks.

```{r statepop, fig.align="center", echo=FALSE, fig.cap="State occupancy over time."}
knitr::include_graphics("../../images/state_pop_over_time.png")
```

### Cost-effectiveness

| Strategy          | Cost, $c$ | $\Delta c$ | QALYs, $e$ | $\Delta e$ | ICER |
|-------------------|-----------|------------|------------|------------|------|
| Baseline          |           |            |            |            |      |
| \> 4%             |           |            |            |            |      |
| \> 6%             |           |            |            |            |      |
| \> 1 risk factor  |           |            |            |            |      |
| \> 2 risk factors |           |            |            |            |      |

: Cost-effectiveness statistics

Figure \@ref(fig:ceac) gives the CEAC and 
Figure \@ref(fig:ceplane) gives the CE-plane.

```{r ceac, fig.align="center", echo=FALSE, fig.cap="Cost-effectiveness acceptability curve (CEAC)."}
knitr::include_graphics("../../images/ceac.png")
```

```{r ceplane, fig.align="center", echo=FALSE, fig.cap="Cost-effectiveness plane."}
knitr::include_graphics("../../images/ceplane.png")
```

## Discussion

<!-- shared decision-making -->

Decision making as to whether have the implant is shared between patient and medical professional.
The SCD risk score is a support tool which contributes to the final decision.
An SCD risk score of 6% is treatment recommendation for one patient versus another.
There could be individual preferences that out-weighs the risk score and mean a patient chooses the alternative.

<!-- better QoL -->

Having more robust quality of life data will help to make such decisions.
more embedded PROMs research

Informatics Consult [@Lai2021] automation to scale evidence generation and to accelerate the return of results within clinical time-scales.

<!-- limitation -->

We did not include CRT-D or s-ICD but that can be overcome by sensitivity analyses that show e.g. the main message is that conclusions about the ICER etc will be robust to most departures from base assumptions other than tweaks in utility (which are the part of the model based on the weakest data!)

[@Olivotto2020] The role of new DMARDs and their ability to influence quality of life.
So if both ICD and non-ICD patients experience significant increases in QoL, are these equally distributed?

# Appendix

```{r barplot, fig.align="center", out.width = "80%", echo=FALSE, fig.cap="Year of study entry bar plot."}
knitr::include_graphics("../../images/year_of_entry_barplot.png")
```
